---
title: "Midpoint Deliverable"
author: "Minho Kim, Jaiden York, Vincent Lee, Ainsley Lai"
date: "5/20/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, warning=FALSE}
# [Ainsley Lai] An introduction of the problem domain and dataset (feel 
#free to borrow language you used in your brainstorm assignment)
# [] A paragraph of summary information, citing at least 5 values calculated
#from the data
# [Minho] A table of  aggregated data (e.g., that has been "grouped" 
#by one of your features)
# [Vincent Lee] Three charts that display information from the data 
#(what you visualize is up to you). 

library("knitr")
library("plotly")
library("ggplot2")
library("dplyr")
source("scripts/first_chart_script.R")
source("scripts/second_chart_script.R")
source("scripts/third_chart_script.R")
source("scripts/table_script.R")
source("scripts/summary_script.R")
firearm_df <- read.csv("data/firearm_homicide.csv")
police_df <- read.csv("data/police_killings.csv", stringsAsFactors = FALSE)
gun_violence_df <- read.csv("data/gun_violence_abridged.csv",
                            stringsAsFactors = FALSE)
```

## Introduction

Our report aims to study the prominent trends in gun violence and malignant 
usage in the United States. We chose to focus on this issue as it has been a hot
topic in recent years, both regarding regular citizens and police forces. By 
studying data surrounding the topic, we hope to uncover insights regarding gun 
violence. Datasets were pulled from [Data-world](https://data.world/azel/
gun-deaths-in-america/workspace/file?filename=full_data.csv), 
[Kaggle](https://www.kaggle.com/ericking310/us-gun-violence), and
[MappingPoliceViolence](https://mappingpoliceviolence.org/). 
Furthermore, these datasets compiled their data from sources such as 
FiveThirtyEight, the Gun Violence Archive, and nationwide police departments.

## Summary Info
```{r, echo=FALSE, warning=FALSE, include=FALSE}
summary <- get_summary_info(firearm_df)
# year_highest_deaths
# year_lowest_deaths
# month_highest_deaths
# most_intent
# most_race
get_summary_info(firearm_df)
```
Based on 2012-2014 data found by FiveThirtyEight and the CDC, we can find that:

- The year with the _highest_ deaths was `r summary$year_highest_deaths`
- The year with the _lowest_ deaths was `r summary$year_lowest_deaths`
- The month with the highest deaths was `r month.name[summary$month_highest_deaths]`
- The most common intent for fatal gun usage was `r tolower(summary$most_intent)`
- The race that was the most affected by firearms were `r summary$most_race` people

## Aggregate Table

```{r, echo=FALSE, warning=FALSE}
kable(grouped_table(firearm_df,
                    "race", "education", "BA+", "Some college", "HS/GED"
                    , "Less than HS"),
      col.names = c("Race", "Total", "Bachelors or higher", "Some college",
                    "HS/GED", "Less than HS"))

```
**Aggregate Table** intends to show the relationship between social staus and 
gun violation victims. 

This aggregate table classifies victims by race and education. Both race and
education represent social status of victims. Through this table, our team tries 
to figure out the relationship between the gun violation victims and social 
status. 

White deaths are highest, followed by Black, Hispanic, Asian, and Native American. 
Basically, there are more gun accident victims in a populous race, but the 
proportion of black victims is higher than Hispanics, indicating that black 
people has more possibility got shot from police . There are also more victims
of high school graduates or people who did not like high school in all races 
except Asians. This means that people involved in the lower the social 
status (especially lower education) have more risk.

## Charts

### Chart 1:


```{r, echo=FALSE, warning=FALSE}
impacted_city(police_df)
```

Chart 1 intends to investigate the question of a lack of aggression and how it
relates to the number of cases of police shootings.


A scatter plot was employed for chart 1 because there are two numeric variables 
being investigated. The number of people who did not express any level of 
aggression is being correlated to the number of police shooting cases in this 
chart. A scatter plot helps answer this because scatter plots work to present 
and reveal the correlation between two variables.

**Insights:** Major cities such as Los Angeles, Houston, Phoenix, and Chicago have 
a relatively high number of cases corresponding to a high number of people who 
were not aggressive to the responding law enforcement officers. This reveals 
that there is potentially a common factor in major cities contributing to 
officer shootings in non-threatening or non-aggressive situations. For example,
a lack of adequate training could be a possibility as major cities need to 
employ more officers since the population is more dense in major cities; this
could result in less well-trained officers.Ideally, the trend would show that 
police shootings and a lack of aggression from the subject individual are 
negatively correlated; however, chart 1 presents a positive correlation. 
Furthermore, it appears that most cities outside the aforementioned major 
cities, have a relatively lower amount of cases. This suggests that there may 
be underlying factors associated with the major cities that contribute to higher 
rates of police shootings in non-aggressive situations.

### Chart 2:

```{r, echo=FALSE,message=FALSE, warning=FALSE}
second_chart(gun_violence_df)
```

Chart 2 intends to investigate the relationship between states and the number of
people directly affected by gun violence.

Chart 2 is a stacked bar chart. The chart presents a visual representation of 
the number of people killed and injured in gun involved incidents in each state 
from 2017. A stacked bar chart was employed because of the ease of comparison it
allows for readers when comparing the consequences of gun shootings in different
states. The stacked bars allow for readers to differentiate between injury and 
death. This is significant because often the amount of injuries is overlooked 
in gun incidents and the focus is drawn towards deaths; however, injuries are 
also impactful and should be considered in gun violence research.

**Insights:** Illinois stands out as the most impacted city by gun violence and 
shootings. This may suggest that there is an underlying factor present in 
Illinois that significantly contributes to gun violence. One such factor could 
be the presence of gang violence, as Illinois, and specifically Chicago, is 
notorious for its gang-related violence. In contrast, Wyoming is the least 
impacted city in terms of number of people killed or injured by gun violence. 
This could suggest that there are specific attributes of Wyoming such 
as effective gun regulations and laws or low population density which help limit
the consequences of gun violence.


### Chart 3:

```{r, echo=FALSE, warning=FALSE}
display_third_chart(firearm_df)
```

Chart 3 intends to investigate the different causes or forms of fatal gun usage
such as suicides or homicides.

Chart 3 is a pie chart depicting a breakdown of the 
different types of fatal gun usage from 2013-2018. A pie chart is useful in this 
case because there are multiple categories of fatal gun usage. Chart 3 allows 
for readers to easily observe the distribution of causes when investigating 
different causes of gun fatality. 

**Insights:** The majority of firearm deaths(62.7%) suffered from 2013-2018 was a 
result of suicide. This suggests that perhaps the ability for mentally unhealthy 
individuals to access firearms is much easier than it should be. In addition, 
accidental fatal gun usage makes up 1.63% of the gun fatalities which suggests 
that current firearm training is effective in preventing accidents. 
